---
title: "Notes.Rmd"
author: "Jonah Merriam"
date: "8/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The data in the original excel file has a lot of weird aspects. One of these is spaces in between the columns of data which will be easier to remove in excel. Another thing is that importing the data from excel will create a mix of numeric and character based rows for the data which will have to be swapped around. 

Before reading the data into R I cleaned the excel sheet up so that there were no spaces in between columns and that the column with the name of whatever level was being looked at was named that (ex: phylum)

Cleaned up the data: switch zeros to NA, remove all rows with only NA, standardized the naming of column headers, made all character columns in lowercase only

```{r Working with long format data}
long_ibs<-read_excel(here("Frost IBS Data", "Frost IBS Long form data.xlsx")) %>% clean_names() %>% mutate(name = str_to_lower(name))
 
```

```{r Phylum Long Graph work}
##Using the new phylum long excel sheet

tax_color=c("darkred","firebrick3","firebrick1","hotpink3","maroon2","hotpink1"," indianred2","orangered2","darkgoldenrod2","yellow2","darkolivegreen2","chartreuse2","springgreen3","darkgreen","turquoise3","turquoise1","lightcyan1","lightskyblue1","lightskyblue3","steelblue2","royalblue1","royalblue3","royalblue4","plum1","mediumorchid1","purple1","mediumorchid4","purple4","snow3","snow4","grey33")

ggplot(Phylum_Long, aes(y=phylum_relative_abundance,fill=phylum,x=name))+ geom_bar(stat='identity',color="black",aes(fill=phylum))+    
scale_fill_manual(values=tax_color)+
    labs(title="", x="",y="Relative abundance of reads")+
    theme_bw() +
    theme(legend.position="right",axis.text.x = element_text(angle=45, hjust=1,vjust=1,color="black"))
##this graph is the kind of representation I want for each level but will likely have to set limits on the top twenty most abundant or something like that
```

```{r Class Relative Abundance Graph}
class_long<-read_excel(here("Frost IBS Data", "Class Long .xlsx")) 

ggplot(class_long, aes(y=class_relative_abundance,fill=class,x=name))+ geom_bar(stat='identity',color="black",aes(fill=class))+ scale_color_paletteer_d("awtools::bpalette")+ labs(title="", x="",y="Relative abundance of reads")+theme_bw() + theme(legend.position="right",axis.text.x = element_text(angle=45, hjust=1,vjust=1,color="black"))
##This graph will include all 38 Classes

class_long_shortened<- class_long %>% filter(class_relative_abundance >= 1)
##This will filter out all of the values less than 1% of relative abundance which will give a much simpler looking graph after repeating the first chunk of code with the new shortened class list

class_barchart_shortened<-ggplot(class_long_shortened, aes(y=class_relative_abundance,fill=class,x=name))+ geom_bar(stat='identity',color="black",aes(fill=class))+ scale_color_paletteer_d("awtools::bpalette")+ labs(title="", x="",y="Class Relative abundance above 1%")+ theme_bw() + theme(legend.position="right",axis.text.x = element_text(angle=45, hjust=1,vjust=1,color="black"))
```

```{r Order Graph}
order_long<-read_excel(here("Frost IBS Data", "Order Long.xlsx"))

order_long_shortened<- order_long %>% filter(order_relative_abundance >= 0.5)

order_barchart_shortened<-ggplot(order_long_shortened, aes(y=order_relative_abundance,fill=order,x=name))+ geom_bar(stat='identity',color="black",aes(fill=order))+ scale_color_paletteer_d("awtools::bpalette")+ labs(title="", x="",y="Order Relative abundance above 0.5%")+theme_bw() + theme(legend.position="right",axis.text.x = element_text(angle=45, hjust=1,vjust=1,color="black"))

order_barchart_shortened
```

```{r Family Relative Abundance Barchart}
family_long<-read_excel(here("Frost IBS Data", "Family Long.xlsx"))

family_long_shortened<- family_long %>% filter(family_relative_abundance >= 1)

family_barchart_shortened<-ggplot(family_long_shortened, aes(y=family_relative_abundance,fill=family,x=name))+ geom_bar(stat='identity',color="black",aes(fill=family))+ scale_color_paletteer_d("awtools::bpalette")+ labs(title="", x="",y="Family Relative abundance above 1%")+theme_bw() + theme(legend.position="right",axis.text.x = element_text(angle=45, hjust=1,vjust=1,color="black"))

family_barchart_shortened
```

```{r Genus Relative Abundance Barchart}
genus_long<-read_excel(here("Frost IBS Data", "Genus Long Alphabetical .xlsx"))

genus_long_shortened<- genus_long %>% filter(genus_relative_abundance >= 1)

genus_barchart_shortened<-ggplot(genus_long_shortened, aes(y=genus_relative_abundance,fill=genus,x=name))+ geom_bar(stat='identity',color="black",aes(fill=genus))+ scale_color_paletteer_d("awtools::bpalette")+ labs(title="", x="",y="Genus Relative abundance above 1%")+theme_bw() + theme(legend.position="right",axis.text.x = element_text(angle=45, hjust=1,vjust=1,color="black"))

genus_barchart_shortened
```

```{r Species Relative Abundance Barchart}
species_long<-read_excel(here("Frost IBS Data", "Species Long Alphabetical.xlsx"))

species_long_shortened<- species_long %>% filter(species_relative_abundance >= 1)

species_barchart_shortened<-ggplot(species_long_shortened, aes(y=species_relative_abundance,fill=species,x=name))+ geom_bar(stat='identity',color="black",aes(fill=species))+ scale_color_paletteer_d("awtools::bpalette")+ labs(title="", x="",y="Species Relative abundance above 1%")+theme_bw() + theme(legend.position="right",axis.text.x = element_text(angle=45, hjust=1,vjust=1,color="black"))

species_barchart_shortened
```



